An optimization of bidirectional update language based on the idea of reversible computing \thanks{Supported by organization x.}

### 2019-09-05 paper/main.tex
update tex 0509

\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
\newcommand{\product}[2]{#1 \ prod \ #2}
\newcommand{\tuple}[2]{(#1 :: #2)}
\newcommand{\rearrS}[2]{rearrS \ #1 \ #2}
\newcommand{\rearrV}[2]{rearrV \ #1 \ #2}
\newcommand{\casebx}[6]{case [#1][#2] \to #3, [#4][#5] \to #6}

@@

% Evaluation result: Comparison with OCaml implementation of miniBiGUL

@@

\begin{itemize}
\item our target language is miniBiGUL (BiGUL minus adaptive case) : very well-behaved
\item Basic explanation of bidirectional transformation
  \begin{itemize}
  \item get = s $\to$ v
  \item put = s * v $\to$ s
  \item explanation about get and put.
  \item We assume putget and getput hold (explanation about putget and getput)
  \item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
  \end{itemize}
\item explanation about put-based BX language: BiGUL
\end{itemize}

\subsection{Semantics of miniBiGUL}

@@

$\putbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
    \tab \product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}$

$\putbx{\rearrS{f}{bx}}{s}{v} = f^{-1}(\putbx{bx}{(f \ s)}{v})$

$\putbx{\rearrV{g}{bx}}{s}{v} = \putbx{bx}{s}{(g \ v)}$

$\putbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} = \\
    \tab \text{if } {cond_{sv_1} \ s \ v} \\
    \tab \text{then let } s' = \putbx{bx_1}{s}{v} \\
    \tab \text{else let } s' = \putbx{bx_2}{s}{v} \\
    \tab \text{fi } cond_{s_1} \ s'; \ \text{return} \ s'$

@@

$\getbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
    \tab \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}$

$\getbx{\rearrS{f}{bx}}{s} = \getbx{bx}{(f \ s)}$

$\getbx{\rearrV{g}{bx}}{s} = g^{-1} (\getbx{bx}{s})$

$\getbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
    \tab \text{if} \ {cond_{s_1} \ s}\\
    \tab \text{then let} \ v' = \getbx{bx_1}{s}\\
    \tab \text{else let} \ v' = \getbx{bx_2}{s}\\
    \tab \text{fi} \ {cond_{sv_1} \ s \ v'}; \ \text{return} \ v'$

@@

\begin{itemize}
\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
\item put$_{REV}$ = s * v $\to$ s * g
\item get$_{REV}$ = s $\to$ v * g
\item introduce (put$_{REV}^{-1}$: s * g $\to$ s * v) and (get$_{REV}^{-1}$:v * g $\to$ s)
\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
\end{itemize}

@@

$\putrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx_1}{s_1}{v_1} \ \text{in}\\
    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_2}{v_2} \ \text{in}\\
        \qtab ((s_1, s_2), (g_1, g_2))$

$\putrev{\rearrS{f}{bx}}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{(f \ s)}{v} \ \text{in}\\
        \qtab (f^{-1} s_1, g_1)$

$\putrev{\rearrV{g}{bx}}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{s}{(g \ v)} \ \text{in}\\
        \qtab (s_1, g_1)$

$\putrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} =\\
    \tab \text{if} \ cond_{sv_1} \ s \ v\\
    \tab \text{then let} \ (s', g) = \putrev{bx_1}{s}{v} \ \text{in}\\
    \tab \text{else let} \ (s', g) = \putrev{bx_2}{s}{v} \ \text{in}\\
    \tab \text{fi} \ cond_{s_1} \ s'; \ \text{return} \ (s', g)$

$\putrev{bx_1 \circ bx_2}{s}{v} =\\
    \tab \text{let} \ (s_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_1}{v} \ \text{in}\\
    \tab \text{let} \ (s_3, g_3) = \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
        \qtab (s_3, (g_1, g_2, g_3))$

@@

$\getrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s_1} \ \text{in}\\
    \tab \text{let} \ (v_2, g_2) = \getrev{bx_2}{s_2} \ \text{in}\\
        \qtab ((v_1, v_2), (g_1, g_2))$

$\getrev{\rearrS{f}{bx}}{s}{v} =\\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{(f \ s)} \ \text{in}\\
        \qtab (v_1, g_1)$

$\getrev{\rearrV{g}{bx}}{s}{v} =\\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{s} \ \text{in}\\
        \qtab (g^{-1} v_1, g_1)$

$\getrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
    \tab \text{if} \ cond_{s_1} \ s\\
    \tab \text{then let} \ (v', g) = \getrev{bx_1}{s} \ \text{in}\\
    \tab \text{else let} \ (v', g) = \getrev{bx_2}{s} \ \text{in}\\
    \tab \text{fi} \ cond_{sv_1} \ s \ v'; \ \text{return} \ (v', g)$

$\getrev{bx_1 \circ bx_2}{s}{v} =\\
    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
    \tab \text{let} \ (v_2, g_2) = \getrev{bx_1}{s} \ \text{in}\\
        \qtab (v_2, (g_1, g_2))$

@@

For the approaches so far, when evaluating a composition of BX programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation. This is a kind of information loss. One question is whether it is possible to calculate the above programs without losing information. And the answer is yes. It comes from the the idea of reversible computation where all information during the evaluation need to be kept. In miniBiGUL, we can do that by tupling $put$ and $get$. The such pair is accepted as input of a function named $pg$ to produce a new one that contains the actual result of a BX program.

@@

$\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
    \tab = (\product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}, \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}) \\
    \tab = \text{let} \ (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \ \text{in} \\
        \qtab \text{let} \ (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \ \text{in} \\
        \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$

$\pg{\rearrS{f}{bx}}{s}{v} \\
    \tab = (f^{-1}(\putbx{bx}{(f \ s)}{v}), \getbx{bx}{(f \ s)}) \\
    \tab = \text{let} \ (s, v) = \pg{bx}{f \ s}{v} \ \text{in} \\
        \qtab (f^{-1} s, v)$

$\pg{\rearrV{g}{bx}}{s}{v} \\
    \tab = (\putbx{bx}{s}{(g v)}, g^{-1} (\getbx{bx}{s})) \\
    \tab = \text{let} \ (s, v) = \pg{bx}{s}{g \ v} \ \text{in} \\
        \qtab (s, g^{-1} v)$\\

Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrS{f}{bx}]\!]$ and $[\![\rearrV{g}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\

$\pg{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} \\
    \tab = (\text{if} \ {cond_{sv_1} \ s \ v} \\
    \qtab \text{then let} \ s' = \putbx{bx_1}{s}{v} \\
    \qtab \text{else let} \ s' = \putbx{bx_2}{s}{v} \\
    \qtab \text{fi} \ cond_{s_1} \ s'; \text{ return } s', \\
    \qtab \quad \text{if} \ {cond_{s_1} \ s}\\
    \qtab \quad \text{then let} \ v' = \getbx{bx_1}{s}\\
    \qtab \quad \text{else let} \ v' = \getbx{bx_2}{s}\\
    \qtab \quad \text{fi} \ {cond_{sv_1} \ s \ v'}; \text{ return } v') \\
    \tab = \text{if} \ {cond_{sv_1} \ s \ v} \ and \ {cond_{s_1} \ s} \\
        \qtab \text{then let} \ (s', v') = \pg{bx_1}{s}{v}\\
        \qtab \text{else let} \ (s', v') = \pg{bx_2}{s}{v}\\
        \qtab \text{fi} \ cond_{v_1} \ s' \ and \ cond_{sv_1} \ s \ v'; \text{ return } (s',v')$\\

A restriction for $\pginline{case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified. The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{case}$ uses above. \\

$\pg{bx_1 \circ bx_2}{s}{v} \\
    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tab = \text{let} \ a = \getbx{bx_1}{s} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
        \qtab (d, c) \\
    \tab = \text{let} \ (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
        \qtab (d, c) \\
    \tab = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
        \qtab \text{let} \ (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \ \text{in} \\
        \qtab (d, c)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $(\text{let } a = \getbx{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$. It is possible if we consider $(\getbx{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid. The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $e$ in the result pair $(d,e)$ of the new let-expression equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

\begin{itemize}
\item In pg approach we always need to supply source and view.
\item If we would like to evaluate get, we do not have any view.
\item For this, we require the programmer to write a function to produce view from source. (This can be the same with get??)
\item Example program: \textcolor{red}{Good example?}
  Evaluate: $\putbx{bx}{s}{v}$ 
  Evalaute: $\getbx{bx}{s}$
  $\pg{bx}{s}{v} = $
\end{itemize}

>>>>

\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
\newcommand{\product}[2]{#1 \ prod \ #2}
\newcommand{\tuple}[2]{(#1 :: #2)}
\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}

@@

% Evaluation result: Comparison with OCaml implementation of minBiGUL

@@

\begin{itemize}
\item our target language is minBiGUL (BiGUL minus adaptive case) : very well-behaved
\item Basic explanation of bidirectional transformation
  \begin{itemize}
  \item get = s $\to$ v
  \item put = s * v $\to$ s
  \item explanation about get and put.
  \item We assume putget and getput hold (explanation about putget and getput)
  \item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
  \end{itemize}
\item explanation about put-based BX language: BiGUL
\end{itemize}

\subsection{Semantics of minBiGUL}

@@

$\putbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
    \tab (\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2})$

$\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$

$\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$

$\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
    \tab \text{if} \ {cond_{sv} \ s \ v} \\
    \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
    \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$

@@

$\getbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
    \tab (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})$

$\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$

$\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$

$\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
    \tab \text{if} \ {cond_{s} \ s}\\
    \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
    \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
    \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$

@@

\begin{itemize}
\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
\item put$_{REV}$ = s * v $\to$ s * gb
\item get$_{REV}$ = s $\to$ v * gb
\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
\end{itemize}

@@

$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
        \qtab (s_1 :: s_2, gb_1 :: gb_2)$

$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
        \qtab (f_2 \ s_1, gb_1)$

$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
        \qtab (s_1, gb_1)$

$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
    \tab \text{if} \ cond_{sv} \ s \ v\\
    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$

$\putrev{bx_1 \circ bx_2}{s}{v} =\\
    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$

@@

$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
        \qtab (v_1 :: v_2, gb_1 :: gb_2)$

$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
        \qtab (v_1, gb_1)$

$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
        \qtab (g_2 \ v_1, gb_1)$

$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
    \tab \text{if} \ cond_{s} \ s\\
    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$

$\getrev{bx_1 \circ bx_2}{s}{v} =\\
    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
        \qtab (v_2, gb_1 :: gb_2)$

@@

For the approaches so far, when evaluating a composition of BX programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation. This is a kind of information loss. One question is whether it is possible to calculate the above programs without losing information. And the answer is yes. It comes from the the idea of reversible computation where all information during the evaluation need to be kept. In minBiGUL, we can do that by tupling $put$ and $get$. The such pair is accepted as input of a function named $pg$ to produce a new one that contains the actual result of a BX program.

@@

$\pg{\product{bx_1}{bx_2}}{s_1 :: s_2}{v_1 :: v_2} \\
    \tab = ((\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})) \\
    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
        \qtab \tab (s_1 :: s_2, v_1 :: v_2)$

$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
        \qtab \tab (f_2 \ s, v)$

$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \qtab \tab (s, g_2 \ v)$\\

Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrs{f_1}{f_2}{bx}]\!]$ and $[\![\rearrv{g_1}{g_2}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\

$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
    \qtab \quad \text{if} \ {cond_{s} \ s}\\
    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\

A restriction for $\pginline{Case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified. The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{Case}$ uses above. \\

\textcolor{blue}{(should substitute a, b, c, d, e, f by s1, s2, v1, v2...)}\\
$\pg{bx_1 \circ bx_2}{s}{v} \\
    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tab = a \Leftarrow \getbx{bx_1}{s}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
            \qtab \tab (d, c) \\
    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
            \qtab \tab (d, c) \\
    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{f}{b}; \\
            \qtab \tab (d, c)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $a \Leftarrow \getbx{bx_1}{s}$ into $(f, a) \Leftarrow \pg{bx_1}{s}{dummy}$. It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid. The last equality substitutes $s$ in $(d, e) \Leftarrow \pg{bx_1}{s}{b}$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $e$ in the result pair $(d,e)$ equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

To better understand about $pg$, let's take a look at some examples. The following is the definition of $phead$ in BiGUL [?]:

    $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
    \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
        \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
            \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$

The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$. $\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$. $\getbx{phead}{s_0}$ returns the head of $s_0$. For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\

In minBiGUL, we can express $phead$ as follows:

    $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
        \qtab f_1 = \lambda (s:ss).(s,ss)\\
        \qtab f_2 = \lambda (s,ss).(s:ss)\\
        \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
            \qtab \qtab g_1 = \lambda v.(v,())\\
            \qtab \qtab g_2 = \lambda (v,()).v\\
            \qtab \qtab bx_v = Replace \ prod \ (Skip (const ()))$\\

Now, we evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:

$\putbx{phead}{[1,2,3]}{100}\\
    \tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
    \tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
    \tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
    \tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
    \tab = f_2 \ (\putbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{(100,())})\\
    \tab = f_2 \ ((\putbx{Replace}{1}{100}) :: (\putbx{(Skip (const ()))}{[2,3]}{()}))\\
    \tab = f_2 \ (100 :: [2,3])\\
    \tab = f_2 \ (100, [2,3])\\
    \tab = [100,2,3]$\\

$\getbx{phead}{[1,2,3]}\\
    \tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
    \tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
    \tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
    \tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
    \tab = g_2 \ (\getbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])})\\
    \tab = g_2 \ ((\getbx{Replace}{1}) :: (\getbx{(Skip (const ()))}{[2,3]}))\\
    \tab = g_2 \ (1 :: ())\\
    \tab = g_2 \ (1,())\\
    \tab = 1$\\

By the definition of $pg$, we have:
$\pg{phead}{[1,2,3]}{100}\\
    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
    \tab = ([100,2,3], 1)$\\

On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
$\pg{phead}{[1,2,3]}{100}\\
    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \{\\
        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
        \qtab \tab (s, g_2 \ v)\\
        \qtab \};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \{\\
        \qtab \tab (s, v) \Leftarrow \pg{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{((100,()))};\\
        \qtab \tab (s, g_2 \ v)\\
        \qtab \};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \{\\
        \qtab \tab (s, v) \Leftarrow \{\\
            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip (const ()))}{[2,3]}{()};\\
            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
        \qtab \tab \};\\
        \qtab \tab (s, g_2 \ v)\\
        \qtab \};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \{\\
        \qtab \tab (s, v) \Leftarrow \{\\
            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
        \qtab \tab \};\\
        \qtab \tab (s, g_2 \ v)\\
        \qtab \};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow \{\\
        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
        \qtab \tab (s, g_2 \ v)\\
        \qtab \};\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
        \qtab \tab (f_2 \ s, v)\\
    \tab = ([100,2,3], 1)$\\

If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:

    \tab $\putbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]}{100} = [[100,2,3],[\ ],[4,5]]$\\
    \tab $\getbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]} = 1$\\

We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:

$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}\\
    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}\\
    \tab = \getbx{phead_2}{[1,2,3]}\\
    \tab = 1$\\

$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}{100}\\
    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}{100})}\\
    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{[100,2,3]}\\
    \tab = [[100,2,3],[\ ],[4,5]]$\\

We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
